---
title: "Map of Ariannas in Milan"
output:
  html_document:
    df_print: paged
---


loading libraries data from tsData.csv


```{r}
library(ggplot2)
library(forcats)
library(data.table)
library(fasttime)
library(raster)
library(extraoperators)
library(TTR)
library(fda)
library(zoo)
library(lubridate)
library(forecast)
library(tseries)
#setting parent directory as working directory

currentDir=getwd()

parentDir=dirname(currentDir)

setwd(parentDir)

#reading data

wiseairdata=fread("data/rawdata.csv", drop = 16)

# converting dates and time to POSIXct class (very fast)

wiseairdata$created_at = fastPOSIXct(wiseairdata$created_at)

# converting pot_id variable into a factor  (fast)
wiseairdata$pot_id = as.factor(wiseairdata$pot_id)


```

selecting and saving the time series of PM2.5 of pot 1091.
Plotting the time series and evaluating how many observations are present

```{r}

pot91=wiseairdata[pot_id==1091 ,]

pot91pm2p5plot=ggplot(pot91) + geom_line(aes(x=created_at,y=pm2p5SPS) )

pot91pm2p5plot


print(paste("observations: ",dim(pot91)[1]))

```

1) The time series has time irregularity.
2) the time series appears non-stationary.
In particular, the noisiness of the time series appears to increase when higher values are reached.
Also, the high number of observations and non-stationarity makes analysis with stationary models non-viable.

To apply SARIMA type models, the time series could be cut in different stages.

For example, the time series appears stationary in august and september.

plotting the time series in this period, hilighting weekend days

```{r}


pot91agosept=pot91[created_at %gele% c("2020-08-15","2020-09-07"),]

pot91pm2p5plot2=ggplot(data=pot91agosept) +
  geom_line(aes(x=created_at,y=pm2p5SPS)) + 
  geom_vline(xintercept=ymd_hms(c("2020-08-21 00:00:00",
                                  "2020-08-23 00:00:00",
                                  "2020-08-28 00:00:00",
                                  "2020-08-30 00:00:00",
                                  "2020-09-04 00:00:00",
                                  "2020-09-06 00:00:00",
                                  "2020-09-11 00:00:00",
                                  "2020-09-13 00:00:00"
                                  )),size=1,color="red")


pot91pm2p5plot2

print(paste("observations: ",dim(pot91agosept)[1]))

```

zooming on this portion,seasonality seems to appear, both on a daily and on a weekly basis.

we apply an hourly mean to obtain a regular time series

```{r}

#loading hourly data

currentDir=getwd()

parentDir=dirname(currentDir)

setwd(parentDir)

hourdata=fread("data/tsData.csv")

hourdata$created_at = fastPOSIXct(hourdata$created_at)

#plotting hourly data

pot91hour=hourdata[pot_id==1091 & created_at %gele% c("2020-08-15","2020-09-07"),]

pot91pm2p5plot3=ggplot(data=pot91hour) +
  geom_line(aes(x=created_at,y=pm2p5SPS)) + 
  geom_vline(xintercept=ymd_hms(c("2020-08-21 00:00:00",
                                  "2020-08-23 00:00:00",
                                  "2020-08-28 00:00:00",
                                  "2020-08-30 00:00:00",
                                  "2020-09-04 00:00:00",
                                  "2020-09-06 00:00:00",
                                  "2020-09-11 00:00:00",
                                  "2020-09-13 00:00:00"
                                  )),size=1,color="red")


pot91pm2p5plot3

print(paste("observations: ",dim(pot91hour)[1]))

# approximating NAs linearly


pot91hourSeries=zoo(pot91hour$pm2p5SPS)

pot91hourSeries=na.approx(pot91hourSeries)

summary(pot91hourSeries)

#plotting ACF, PACF

plot(pot91hourSeries)


acf(pot91hourSeries, lag.max=1000)


pacf(pot91hourSeries, lag.max=300); abline(v=168,col="red")

adf.test(pot91hourSeries,k=0)

pot91model=auto.arima(pot91hourSeries)

summary(pot91model)

plot(forecast(pot91model,h=48))


```







